Internet of Things: a Real-World Deployment Towards Smart Grids Integration

### 2013-07-03 doc/thesis/report.tex

\documentclass[oneside,12pt,a4paper,final]{book}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{arabtex}
\usepackage{utf8}
\newcommand{\HRule}{\rule{\linewidth}{0.6mm}}
%\usepackage{hyperref}
\usepackage{listings}
\lstset{language=C}
\usepackage[acronym]{glossaries} 
\makeglossaries
\author{Nacer Khalil}
\title{Internet of Things: a Real-World Deployment Towards Smart Grids Integration}

@@

\chapter{Findings}

\section{Experiment}
%\subsection{Methodology}
\section{Data Gathering}
\section{Interpretation of Results}

@@

\chapter{Future Work}
\paragraph{}
This thesis project's was to establish the infrastructure for integrating smart home within \glsreset{sg}\gls{sg} and make it part of the \glsreset{iot}\gls{iot}. This project enabled the deployment of the two-way communication as an infrastructure that can be used to build the \glsreset{dr}\gls{dr}. \gls{dr} is one of the most important topics in smart grids. It enables the utility and households to communicate with a set of services that any of them may request from the other. The current state of the art does not incorporate this communication as part of the smart home systems \cite{ref15}. But as there is a need for this communication, \gls{dr} will have to be implementation as part of the

\backmatter

\appendix
\chapter{How to install the system}
\section{Hardware needed}
In order for the user to install the system, one should make sure he/she is having a computer that will be used for the sole purpose of hosting different services. The crossbow professional kit is a necessity and a router must be available to connect the different components through Wi-Fi. A smartphone or a separate computer with Android emulator may de used.

@@

\singlespacing
\bibliographystyle{IEEEtran} 
\bibliography{IEEEabrv,myrefs}

>>>>

\documentclass[oneside,12pt,a4paper,final]{book}
%\usepackage[margin=3cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{arabtex}
\usepackage{utf8}
\newcommand{\HRule}{\rule{\linewidth}{0.6mm}}
%\usepackage{hyperref}
\usepackage{listings}
\lstset{language=C}
\usepackage[acronym]{glossaries} 
\makeglossaries
\author{Nacer Khalil}
\title{Internet of Things: a Real-World Deployment Towards Smart Grids Integration}

@@

\chapter{Findings}
\paragraph{}
The system was built to provide monitor and control home appliances and also provide an infrastructure for future platforms to be integrated in this system. Such infrastructures can be the integration of the \glsreset{dr}\gls{dr}. To build overlying system's, one should make sure that the system has a certain level of performance that one can count on to build overlying services. The purpose of this chapter is to measure the performance of the system. It is done by first doing and experiment which will help us gather data and then based on this data, an interpretation of this data will be done to see whether the system is shows an acceptable level of performance.
\section{Experiment}
\paragraph{}
To measure the performance of the system, an experiment was conducted where the goal is to quantify the performance of the system. \gls{6lowpan} was built in order for the mote receiving the packet to forward it is not directed to it before dealing with the one directed to it. This means that there is more priority given to routed packet that the processed ones. As a result, this clearly pushes us to think that the delay of the system will not be much affected as the traffic in the network grows.
\paragraph{Experiment 1:}
We will first have one mote in the \gls{wsn} that will be interacting with the outside world. We will start sending ping command periodically (once every 1 second) and record the delay and the jitter. Once every two minutes, we will add another mote to the system and continue pinging the same mote. We keep adding motes every 2 minutes until we fill the \gls{wsn} with 6 motes which is the maximum number of motes available in the system. This will enable us to collect data about the mote delay and jitter as we add more traffic to the system.
\paragraph{Experiment 2: }
The second experience is meant to measure the time it takes to transform the packet. In other words, measure the average delay and jitter that is added when having the Gateway Transformation Process as part of the system. The timer starts before the sniffing of the packet and ends right after the packet is transformed and sent.
We believe that there will be an additional delay to the communication but we are hoping that it will not affect much the whole delay of the communication between different components of the system. 
\section{Data Gathering}
\paragraph{}
After both experiments were conducted, we collected a significant amount of data that was used to measure the performance and an interpretation was done in section 6.3 to have an idea about the performance of the system.
\subsection{Experiment 1}
\paragraph{}
In the first experiment, the same mote was pinged once every second and we added new motes to the system every two minutes. The experiment as a whole took 12 minutes and 720 pings we done. Table \ref{table:exp1} shows the result of the experiment.

\begin{table}[htbp]
    \begin{tabular}{lll}
    \hline
    number of motes & Delay (ms) & Jitter (ms) \\\hline
    1               & 87.5         & 1.2         \\ 
    2               & 88.1         & 2.0         \\
    3               & 88.3        & 2.7         \\
    4               & 88.8         & 3.1         \\
    5               & 90.0         & 3.3         \\
    6               & 89.5         & 3.5         \\
    \end{tabular}
    \caption{Experiment 1 delay and jitter}
    \label{table:exp1}
\end{table}
\paragraph{}
For each row in the table, the average delay is computed using the pings that were performed in the 2 minutes interval which means 120 pings per row. The jitter is also computed in a similar as the average variation from the mean delay in that 2 minutes interval. The result is displayed in the table \ref{table:exp1}.

\subsection{Experiment 2}
\paragraph{}
The second experiment's goal is to measure the additional delay coming from the Gateway Packet transformation process whose goal to handle and hide the heterogeneity of the system. The result gathered reflects the average delay and jitter computed over elapsed time from the sniffing of the packet to the transformation and sending to the recipient. Table \ref{table:exp2} shows the result that was collected from this experiment as it was done on 200 packets that were sniffed and transformed by the system.

\begin{table}[htbp]
    \begin{tabular}{ll}
    \hline
    Delay (ms) & Jitter (ms) \\ \hline
    0.1        & 0.0         \\ 
    \end{tabular}
    \caption{Experiment 2 delay and jitter}
        \label{table:exp2}
\end{table}
The transformation process's elapsed time is counted in nanoseconds and from \ref{table:exp2}, we can see that the average delay is on average 0.1 milliseconds whereas the jitter is 0 milliseconds meaning that the process's time varies between a few nanoseconds to at most 150 nanoseconds. Depending on the machine's load.
\section{Interpretation of Results}
\subsection{Experiment 1}
The data collected in section 6.2 shows that the performance of the system is very good. Even though the delay in the \gls{wsn} is high, 
 nothing can be done about that as even though the traffic in the \gls{wsn} is low, the delay is still high and this has to do with the computational time taken by the TCP/IP protocol protocol stack provided implemented by \glsreset{blip}\gls{blip} using \glsreset{6lowpan}\gls{6lowpan}. To reduce the delay, one should probably rethink the design of the \gls{blip} or find a way to omit it, but that is not a good idea since we want the \gls{wsn} to be part of \gls{iot}. 
 The good news in experiment 1 is that the privatization of packets brings benefit to the network. Even though the traffic grows exponentially, the delay does not grow much and the even if the jitter follows to some extent the traffic's trend but still the average delay shows good result. 
 \subsection{Experiment 2}
 The Packet transformation process is a component that shows excellent results. Before conducting experiment 2, we were persuaded that there will be additional delay originating from this process and it could be responsible for the high delay seen in experiment 2 but the results shows otherwise and that the component is not to be put responsible for such important delay seen in experiment 1.

@@

This project enabled us to establish an infrastructure as part of the \glsreset{iot}\gls{iot} and move towards having the smart home system's integrated within the smart. At this stage and counting solely on this work, one cannot clearly see the integration of smart home systems within the \gls{iot}, but this is an infrastructure that provides two main services which are the establishment of the two-way communication between the smart home devices and the outside world which may be the utility for instance. This two-way communication enables the \glsreset{wsn}\gls{wsn} to monitor home appliances and send sensor data to a repository which monitors the home's consumption and this device can be used for simulating a real world \gls{ami}. The other part of the communication is the ability of clients to control appliances status by turning them On and Off.
The second contribution of this work the smart management of the heterogeneity. The communication from the mote to the mobile client is built on a set of technologies and the journey of a simple \gls{tcp} datagram move through many different data link layer standards and different network technologies. Starting from the mote sending sensor data, The TCP data is encapsulated into an \gls{ipv6} packet that is encapsulated into a Zigbee frame. Afterward, the packet is extracted from the Zigbee frame and encapsulated into a \gls{usb} frame. Once it arrives to the gateway, the \gls{tcp} datagram is extracted from the \gls{ipv6} which is extracted from the \gls{usb} frame and encapsulated in an \gls{ipv4} packet that encapsulated into a Wi-Fi frame and sent until it reaches the mobile. That is said, it clear that the system is heterogeneous and making this heterogeneity transparent was a contribution of this work.
The \glsreset{wsn}\gls{wsn} was built making use of the very recent \glsreset{6lowpan}\gls{6lowpan} whose goal is to provide the TCP/IP protocol stack to the smallest devices. The \gls{wsn} is autonomous in the sense that it can work independently and indefinitely by monitoring the home appliances. 
All in all, this project as a big picture is a testbed providing the main components of a smart home management with an intelligent \gls{wsn}, a robust and powerful middleware application that is able to use the sensory data and use it to control the home's appliances and a mobile application able to provide the users remote monitoring of their smart homes and control. This was possible by making the system as part of the \glsreset{iot}\gls{iot}.
\chapter{Future Work}
This thesis project's was to establish the infrastructure for integrating smart home within \glsreset{sg}\gls{sg} and make it part of the \glsreset{iot}\gls{iot}. This project enabled the deployment of the two-way communication as an infrastructure that can be used to build the \glsreset{dr}\gls{dr}. \gls{dr} is one of the most important topics in smart grids. It enables the utility and households to communicate with a set of services that any of them may request from the other. The current state of the art does not incorporate this communication as part of the smart home systems \cite{ref15}. But as there is a need for this communication, \gls{dr} will have to be implemented as part of the smart home package.
Demand reponse integration is a future work for this study. \gls{dr} has a set of service that have to be implemented in the smart home packages. In \gls{dr}, the consumer has the possibility to ask the utility about the instantaneous price quote for electricity. The consumer when in need of high priority in energy may ask the utility to provide such to provide a minimum of energy that will be guaranteed. Of course this means at a cost, but this agreement is part of the \gls{dr} and should be implemented at the level of smart home systems. Consumers can also help utilities at difficult times. For instance, when the utility cannot provide enough energy at peak times, the utility may ask some consumers to significantly cut their consumption and as a result be billed a reduced cost for the electricity that is consumed. In general, the smart home aims to reduce energy consumption and as the utility at times may need to reduce its energy production especially in peak hours, the \gls{dr} will help the utility ask the smart home to reduce the home consumption.
As this project aims to integrate the smart home devices within the \glsreset{iot}\gls{iot}, there is a need to integrate \glsreset{rfid}\gls{rfid} within this system as a future work. The \gls{rfid} is a building block within the \gls{iot} and will equip the devices with a low-power radio identification that will allow more flexibility to the system, as the external systems may be able to recognize uniquely each device in the \gls{iot} without counting primarily on the \gls{ipv6} which may change from time to time and therefore cannot keep track of the each device's information and monitor it reliably.
This system provides near real-time sensory data that is not much exploited intelligently. The only use of this data is to enable users to monitor their home consumption. As far as the system is concerned, there is a need for a more intelligent exploitation of this very rich data that can be used by intelligent algorithm to automatically and autonomously reduce the household's energy consumption. Such algorithms can be based on machine learning and data mining techniques making the direction of this a cross disciplinary project aiming to reduce the consumption of household without reducing the utility of users. In other words, build algorithms that will reduce consumption of households while the user is not noticing change.
Before building new services on top of this system, one should test different aspects of the system. In this study, the only thing that was tested, measured and quantified was the performance of the system which gave good results about the system's overall performance. As a future work, one should test different aspects of the system such as the reliability and availability of the system. More experiments should be conducted to see how well the system acts in high traffic networks and how good could it support new overlying systems such as \gls{dr}.
The system provides different services, but there is not authentication or authorization involved. This means that any users knowing the mote connected to the appliance he/she is interested in, he/she can control it and there is no existing party or component that will control such a communication. As a future work, the control of the system's access should be monitored in order for the system to be more secure and robust to external attacks.

\backmatter
\singlespacing

\addcontentsline{toc}{chapter}{Bibliography}
\renewcommand{\bibname}{References}
\bibliographystyle{plain} 
\bibliography{myrefs}

\doublespacing
\appendix
\chapter{How to install the system}
\section{Hardware needed}
In order for the user to install the system, one should make sure he/she is having a computer that will be used for the sole purpose of hosting different services. The crossbow professional kit is a necessity and a router must be available to connect the different components through Wi-Fi. A smartphone or a separate computer with Android emulator may de used.

@@
